---
title: "Vendor Analysis"
author: "Shathuki Perera"
date: "2024-02-07"
output: html_document
---
# Setting up data
```{r}
# set working directory 
#setwd()
```

```{r, message=FALSE, warning=FALSE}
# Load all necessdary packages the packages 
library(tidyverse)
library(dplyr)
library(ggplot2)
library(tidyr)
library(reshape2)
library(viridis)
library(gridExtra)
library(RColorBrewer)
library(stats)
```

```{r}
# load the data 
data <- read.csv("R data .csv")
```

```{r, results='hide'}
# Formatting the column names
colnames(data)[colnames(data) %in% c("Dry..D..or.Fresh..F..fish", "Biological.sex", "Age", "How.long.have.you.been.selling.fish.in.the.market.", "What.quantity.of.sharks..kg..do.you.sell.per.day.", "What.quantity.of.rays..kg..do.you.sell.per.day.", "Which.species.of.shark.and.ray.do.you.sell...Show.pictures.", "species", "How.much.is.a.kg.of.shark...Rps.", "How.much.is.a.kg.of.ray...Rps.")] <- c("type", "sex", "age", "time.selling", "quantity.shark", "quantity.ray", "species.sold", "species.preferred", "shark.price", "ray.price" )
```

### Comparing the quantities of shark versus ray sold

```{r, results = 'hide'}
# I am using a paired t-test for this
# check the data types
str(data$quantity.shark)
str(data$quantity.ray)
```

```{r, warning=FALSE}
# Convert columns to numeric
data$quantity.shark <- as.numeric(data$quantity.shark)
data$quantity.ray <- as.numeric(data$quantity.ray)
```

```{r}
# Performing an independent t-test, as the data is not paired for all values and some vendors only sold shark, or ray. 
# Create a long format of the data for plotting
data_long <- pivot_longer(data, cols = c(quantity.shark, quantity.ray), names_to = "Species", values_to = "Quantity")
```

```{r}
# Independent samples t-test
t_test_result <- t.test(Quantity ~ Species, data = data_long)

# Print the t-test results
cat("Independent samples t-test results:\n")
print(t_test_result)
```

# Non-parametric test: Wilcoxon signed rank test 
```{r}
# Perform Wilcoxon signed-rank test
wilcox_test_result <- wilcox.test(data$quantity.shark, data$quantity.ray, paired = TRUE, na.rm = TRUE)

# Print the result
print(wilcox_test_result)
```
```{r}
# Create a boxplot for visualisation
ggplot(data_long, aes(x = Species, y = Quantity, fill = Species)) +
  geom_boxplot() +
  labs(title = "Comparison of Shark and Ray Quantities",
       x = "Species",
       y = "Quantity") +
  scale_fill_manual(values = c("quantity.shark" = "steelblue", "quantity.ray" = "coral")) +
  theme_minimal()

```

```{r}
# Removing the outliers
ggplot(data_long, aes(x = Species, y = Quantity, fill = Species)) +
  geom_boxplot() +
  labs(title = "Comparison of Shark and Ray Quantities",
       x = "Species",
       y = "Quantity") +
  scale_fill_manual(values = c("quantity.shark" = "steelblue", "quantity.ray" = "coral")) +
  theme_minimal() +
  coord_cartesian(ylim = c(0, 300))

```

Interpretation:
The negative t-value (-0.52524) suggests that, on average, there is a slightly lower quantity of shark compared to ray, but the magnitude is not very large.
We do not have sufficient evidence to conclude that there is a significant difference in the quantities of shark and ray sold by vendors, and the p-value is greater than the commonly used significance level of 0.05. Thus, based on the t-test results, we do not have strong evidence to suggest a significant difference between the quantity of shark and ray sold. 
The point is also proven by the Wilcoxon signed rank test- not significant. 

There are also some obvious anomalies of data points which are very big. This is likely to be due to the differences between the retail vendors versus wholesale vendors. 

```{r}
# Remove rows with NAs in either quantity.shark or quantity.ray
data_clean <- na.omit(data[c("quantity.shark", "quantity.ray")])

# Calculate summary statistics for Shark and Ray separately
summary_stats <- data_clean %>%
  summarize(
    mean_quantity_shark = mean(quantity.shark),
    max_quantity_shark = max(quantity.shark, na.rm = TRUE),
    min_quantity_shark = min(quantity.shark, na.rm = TRUE),
    sd_quantity_shark = sd(quantity.shark, na.rm = TRUE),
    mean_quantity_ray = mean(quantity.ray),
    max_quantity_ray = max(quantity.ray, na.rm = TRUE),
    min_quantity_ray = min(quantity.ray, na.rm = TRUE),
    sd_quantity_ray = sd(quantity.ray, na.rm = TRUE)
  )

# Print summary statistics
print(summary_stats)
```

### Comparing the quantity of shark versus ray according to their location
```{r}
# Replace NA with 0 for interviewees who didn't sell a particular type
data[is.na(data)] <- 0
```

```{r}
# Create a long format of the data for plotting
data_long2 <- pivot_longer(data, cols = c(quantity.shark, quantity.ray), names_to = "Species", values_to = "Quantity")

# Specify the order of levels for 'Location' and remove unwanted levels
data_long2$Location <- factor(data_long2$Location, levels = c('MH', 'FM', 'K'), labels = c('Fisheries Harbour', 'Fish Market', 'Kamachchode Market'))
```

```{r}
# Create box plots for shark and ray quantities within each location, with the same scale
# Making the scale the same for all plots 
ggplot(data_long2, aes(y = Quantity, fill = Species)) +
  geom_boxplot(position = "dodge") +
  labs(title = "Comparison of Shark and Ray Quantities by Location",
       x = "Location",
       y = "Quantity") +
  scale_fill_manual(values = c("quantity.shark" = "steelblue", "quantity.ray" = "coral")) +
  theme_minimal() +
  facet_wrap(~Location, scales = "free_y")+
  theme(axis.text.x = element_text(size = 7.5))+ 
coord_cartesian(ylim = c(0, 300))
```

```{r}
# Statistical test to check for significance
# Perform pairwise comparisons for each market
pairwise_tests <- data_long2 %>%
  group_by(Location) %>%
  do(pairwise_test = wilcox.test(Quantity ~ Species, data = ., paired = TRUE, exact = FALSE, na.rm = TRUE))

# Print the results
print(pairwise_tests)
```

Interpretation:
Here we can see that most of the anomalies are in the Fisheries Harbour. This makes contexual sense, as this is where most of the wholesale traders were present, thus a greater amount of shark and ray are traded here. Also, vendors at the Negombo Fish Market would travel to the Harbour in the morning to get shark and ray to sell later on the day in the market. 

```{r}
# Test for significance in the quantities of shark and ray based on location
# Perform Wilcoxon signed-rank test
wilcox_test_result2 <- wilcox.test(data$quantity.shark, data$quantity.ray, paired = TRUE, na.rm = TRUE, exact = FALSE)
```

```{r}
#see the results
print(wilcox_test_result2)
```
# Comparing shark and ray quantities according to type, instead of location

```{r}
# Plotting boxplots for shark and ray quantities separated by type
ggplot(data_long, aes(x = Species, y = Quantity, fill = Species)) +
  geom_boxplot() +
  labs(title = "Comparison of Shark and Ray Quantities", x = "Species", y = "Quantity") +
  scale_fill_manual(values = c("steelblue", "coral"), name = "Species", labels = c("Shark", "Ray")) +
  theme_minimal() +
  coord_cartesian(ylim = c(0, 300)) +
  facet_wrap(~ type, labeller = labeller(type = c("D" = "Dry Fish", "F" = "Fresh Fish")))
```

```{r}
# ANOVA for quantity against type
# Perform ANOVA for shark prices across type
anova_shark_quantitytype <- aov(quantity.shark ~ type, data = data)
summary(anova_shark_quantitytype)

# Perform ANOVA for ray prices across type
anova_ray_quantitytype <- aov(quantity.ray ~ type, data = data)
summary(anova_ray_quantitytype)

##No significance of quantity depending on type: either dry or fresh
```

```{r}
# ANOVA for quantity against location
# Perform ANOVA for shark quantity across locations
anova_shark_qlocation <- aov(quantity.shark ~ Location, data = data)
summary(anova_shark_qlocation)

# Perform ANOVA for ray quantity across locations
anova_ray_qlocation <- aov(quantity.ray ~ Location, data = data)
summary(anova_ray_qlocation)
```

```{r}
## comparing the mean and standard deviations of quantity ray and shark between locations:
# Calculate averages and standard deviations
summary_stats <- data_long2 %>%
  group_by(Species) %>%
  summarize(avg_quantity = mean(Quantity),
            sd_quantity = sd(Quantity))
summary_stats
```

### Comparing the prices of shark versus ray 

```{r}
#check data type
str(data$shark.price)
str(data$ray.price)

# change data type
# Convert columns to numeric
data$shark.price <- as.numeric(data$shark.price)
data$ray.price <- as.numeric(data$ray.price)
```

```{r}
# Descriptive statistics
# Calculate mean and standard deviation for shark.price
mean_shark_price <- mean(data$shark.price, na.rm = TRUE)
sd_shark_price <- sd(data$shark.price, na.rm = TRUE)

# Calculate mean and standard deviation for ray.price
mean_ray_price <- mean(data$ray.price, na.rm = TRUE)
sd_ray_price <- sd(data$ray.price, na.rm = TRUE)

# Print the results
cat("Mean Shark Price:", mean_shark_price, "\n")
cat("Standard Deviation Shark Price:", sd_shark_price, "\n")
cat("Mean Ray Price:", mean_ray_price, "\n")
cat("Standard Deviation Ray Price:", sd_ray_price, "\n")

```

```{r}
# Descriptive statistics based on location
# Calculate mean and standard deviation for shark.price and ray.price separately for each location
summary_stats <- data %>%
  group_by(Location) %>%
  summarise(
    mean_shark_price = mean(shark.price, na.rm = TRUE),
    sd_shark_price = sd(shark.price, na.rm = TRUE),
    mean_ray_price = mean(ray.price, na.rm = TRUE),
    sd_ray_price = sd(ray.price, na.rm = TRUE)
  )

# Print the summary statistics
print(summary_stats)
```

```{r}
# Perform ANOVA for shark prices
anova_shark <- aov(shark.price ~ Location, data = data)

# Perform Tukey's HSD test for shark prices
tukey_shark <- TukeyHSD(anova_shark)
print(tukey_shark)

# Perform ANOVA for ray prices
anova_ray <- aov(ray.price ~ Location, data = data)

# Perform Tukey's HSD test for ray prices
tukey_ray <- TukeyHSD(anova_ray)
print(tukey_ray)
```

```{r}
data_long3 <- pivot_longer(data, cols = c(shark.price, ray.price), names_to = "species", values_to = "price")

# Plotting boxplot with ggplot2
ggplot(data_long3, aes(x = species, y = price, fill = species)) +
  geom_boxplot() +
  labs(title = "Boxplot of Shark and Ray Prices", x = "Species", y = "Price") +
  theme_minimal() +
  scale_fill_manual(values = c("steelblue", "coral"))
```
# Looking at Price differences between Shark and Ray, and also between Type
```{r}
# Filter data for 'F' (fresh fish) and 'D' (dry fish) types
data_fresh <- filter(data_long3, type == 'F')
data_dry <- filter(data_long3, type == 'D')

# Combine data for fresh and dry fish
combined_data <- rbind(data_fresh, data_dry)

# Define custom labels for facet_wrap
custom_labels <- c("D" = "Dry Fish", "F" = "Fresh Fish")

# Plotting combined boxplots for fresh and dry fish with custom facet labels
ggplot(combined_data, aes(x = species, y = price, fill = species)) +
  geom_boxplot() +
  labs(title = "Boxplot of Shark and Ray Prices based on Type", x = "Species", y = "Price") +
  theme_minimal() +
  scale_fill_manual(values = c("steelblue", "coral")) +
  facet_wrap(~ type, labeller = labeller(type = custom_labels))
```

```{r}
# Cleaning data set to run statistical test 
# Select only the columns 'type', 'species', and 'price' in combined_data
combined_data <- combined_data %>%
  select(type, species, price)
```

```{r}
# Filter data for shark 
shark_data <- combined_data %>%
  filter(grepl("shark.price", species))
# Running test for shark 
# Perform Mann-Whitney U test
wilcox.test(price ~ type, data = shark_data)

#p= 0.0003605 is significant 
```
```{r}
# Filter data for ray 
ray_data <- combined_data %>%
  filter(grepl("ray.price", species))
# Running test for shark 
# Perform Mann-Whitney U test
wilcox.test(price ~ type, data = ray_data)

#significant p value
```

```{r}
# Calculate the mean price for 'D' type
mean_price_D <- mean(shark_data$price[shark_data$type == 'D'], na.rm = TRUE)

# Calculate the mean price for 'F' type
mean_price_F <- mean(shark_data$price[shark_data$type == 'F'], na.rm = TRUE)

# Print the mean prices
cat("Mean price for 'D' type:", mean_price_D, "\n")
cat("Mean price for 'F' type:", mean_price_F, "\n")

```

```{r}
# Calculate the mean price for 'D' type in ray_data
mean_price_D_ray <- mean(ray_data$price[ray_data$type == 'D'], na.rm = TRUE)

# Calculate the mean price for 'F' type in ray_data
mean_price_F_ray <- mean(ray_data$price[ray_data$type == 'F'], na.rm = TRUE)

# Print the mean prices for ray_data
cat("Mean price for 'D' type in ray_data:", mean_price_D_ray, "\n")
cat("Mean price for 'F' type in ray_data:", mean_price_F_ray, "\n")

```

```{r}
# Testing for normality 
# Example Q-Q plot for shark prices
qqnorm(data$shark.price)
qqline(data$shark.price)

# Shapiro-Wilk test for shark prices
shapiro.test(data$shark.price)

# Example boxplot
boxplot(shark.price ~ ray.price, data = data)
```

```{r}
# To test if the price of ray and shark are significantly different from each other, I have used the Wilcoxon signed-rank test.
# Perform Wilcoxon signed-rank test
wilcox_test_result2 <- wilcox.test(data$shark.price, data$ray.price, paired = TRUE, na.rm = TRUE)

# Print the result
print(wilcox_test_result2)
```
Interpretation:
The extremely small p-value (1.565e-05) suggests strong evidence that there is a significant difference in the price between sharks and rays. 

```{r}
# ANOVA for price against location
# Perform ANOVA for shark price across locations
anova_shark_plocation <- aov(shark.price ~ Location, data = data)
summary(anova_shark_plocation)

# Perform ANOVA for ray price across locations
anova_ray_plocation <- aov(ray.price ~ Location, data = data)
summary(anova_ray_plocation)
```

```{r}
# Load the TukeyHSD function from the stats package
library(stats)

# Perform Tukey's HSD test for shark price across locations
tukey_shark_plocation <- TukeyHSD(anova_shark_plocation)
print(tukey_shark_plocation)

# Perform Tukey's HSD test for ray price across locations
tukey_ray_plocation <- TukeyHSD(anova_ray_plocation)
print(tukey_ray_plocation)

```


### Plot for the species count for all unique species mentioned by vendors
# Question 'What species of shark and ray do you sell?'

```{r}
## insert new data set with species without NAs
species.count <- read.csv("species.count.csv")
```

```{r}
my_palette <- c(
  "Alopiidae" = "#ff7f0e",
  "Lamnidae" = "#9467bd",
  "Carcharhinidae" = "#2ca02c",
  "Dasyatidae" = "#d62728",
  "Mobulidae" = "#8c564b",
  "Rhinobatidae" = "#e377c2",
  "Rhinopteridae" = "#7f7f7f",
  "Aetobatidae" = "#1f77b4",
  "Sphyrnidae" = "#bcbd22"
)
```


```{r}
species.count$Species <- factor(species.count$Species, levels = species.count$Species[order(-species.count$Count)])
```

```{r}
# Create the bar chart using ggplot2 with color based on Type
plot1.0 <- ggplot(species.count, aes(x = Species, y = Count, fill = Type)) +
  geom_bar(stat = "identity") +
  labs(title = "Count of Species: Shark vs. Rays",
       x = "Species",
       y = "Count") +
  scale_fill_manual(values = c("steelblue", "coral")) +  # Specify colors
  theme_minimal() +
  theme(axis.text.x = element_text(size = 6, angle = 45, hjust = 1),
        legend.text = element_text(size = 7),  
        legend.title = element_text(size = 7),
        axis.text.y = element_text(size = 6),
        plot.title = element_text(size = 10))
plot1.0
```
```{r}
# Reorder Species based on Family
species.count$Species <- factor(species.count$Species, levels = unique(species.count$Species[order(species.count$Family)]))

# Plot
plot2.0 <- ggplot(species.count, aes(x = Species, y = Count, fill = Family)) +
  geom_bar(stat = "identity") +
  labs(x = "Species", y = "Count", title = "Count of Species by Family") +
  scale_fill_manual(values = my_palette) +  # Use custom color palette
  theme_minimal() +
  theme(axis.text.x = element_text(size = 6, angle = 45, hjust = 1),
        legend.text = element_text(size = 7),  
        legend.title = element_text(size = 7),  
        axis.text.y = element_text(size = 6),
        plot.title = element_text(size = 10)) +
  guides(fill = guide_legend(title = "Family"))  # Add legend title

plot2.0
```

```{r}
# Arrange plots side by side
grid.arrange(plot1.0, plot2.0, ncol = 2)
```

### Salient Analysis for the species most preferred 
## Frequency of the listed species
# Question 'Which species of shark and ray do you trade the most?'

```{r}
# Because the species are listed, I will separate them and split them into separate rows
data_split <- data %>%
  separate_rows(species.preferred, sep = ", ") #creates new row for each species mentioned
```

```{r}
#cleaning up any leading or trailing white space 
data_split$species.preferred <- trimws(data_split$species.preferred)
```

```{r}
#Calculating the frequency of each species
species_frequency <- data_split %>%
  group_by(species.preferred) %>%
  summarise(frequency = n())
```

```{r}
#Ranking and visualising the data
ranked_species1 <- species_frequency %>%
  arrange(desc(frequency))
```

```{r}
## Trying colours
ranked_species1 <- species_frequency %>%
  arrange(desc(frequency))

# Define a categorical colour palette
colour_palette <- viridisLite::viridis(length(ranked_species1$species.preferred), option = "D")

# Bar plot with colour, smaller and angled x-axis labels
ggplot(ranked_species1, aes(x = species.preferred, y = frequency, fill = species.preferred)) +
  geom_bar(stat = "identity") +
  scale_fill_manual(values = colour_palette) +  # Apply color palette
  labs(title = "Salient Species Analysis",
       x = "Species",
       y = "Frequency") +
  theme(axis.text.x = element_text(size = 8, angle = 45, hjust = 1)) +
  guides(fill = FALSE)  # Remove legend
```

### Calculating salience using equation

```{r}
# Creating new data set for salient analysis
data1 <- data %>%
  select(Interview.Code, species.preferred)
```

```{r}
# Counting the number of interviews before removing N/A 
num_interviews_before_NA <- nrow(data1)
```

```{r}
#Making data1.0 
# Function to split species and add order column
split_species <- function(df) {
  df %>%
    filter(species.preferred != "N/A") %>%
    mutate(species.preferred = strsplit(as.character(species.preferred), ", ")) %>%
    unnest(species.preferred) %>%
    mutate(species.preferred = str_squish(species.preferred)) %>%
    group_by(Interview.Code) %>%
    mutate(order = row_number()) %>%
    ungroup()
}

# Applying the function to your original data
data1.0 <- split_species(data1)
```

```{r}
#Calculating salience from cleaned data
calculate_salience <- function(df) {
  # Calculate salience scores for each species
  salience_scores <- df %>%
    group_by(Interview.Code) %>%
    mutate(salience = (1 + length(species.preferred) - order) / length(species.preferred))

  return(salience_scores)
}

# Apply the function to your data1.0 dataframe
SC <- calculate_salience(data1.0)
```

```{r}
# Adding column to show whether it is a shark or a ray 
# Reorder the data frame based on salience scores
SC <- SC[order(SC$salience, decreasing = TRUE), ]

# Create a new column for color based on the presence of 'shark' in the species name
SC$type <- ifelse(grepl("shark", tolower(SC$species.preferred)), "Sharks", "Rays")

# Display the result       
print(SC)
```

```{r}
# Calculate the mean rank for Sharks vs Rays
mean_rank <- SC %>%
  group_by(type) %>%
  summarize(mean_rank = mean(order))

# Calculate the range of salience scores for Sharks vs Rays
score_range <- SC %>%
  group_by(type) %>%
  summarize(min_salience = min(salience),
            max_salience = max(salience),
            range_salience = max_salience - min_salience)

# Combine the results into a single data frame
summary_stats <- bind_rows(mean_rank, score_range)

# Print the summary statistics
print(summary_stats)
```

```{r}
#Anova test to see if the salience scores are different from Sharks versus Rays
# Fit the ANOVA model
anova_result <- aov(salience ~ type, data = SC)

# Print the ANOVA table
print(summary(anova_result))
```

## Culutural Salience Calculations

```{r}
# Calculate cultural salience for each unique species
cultural_salience_df <- SC %>%
  group_by(species.preferred) %>%
  summarise(CS = sum(salience) / n_distinct(Interview.Code))
```

```{r}
# Calculate cultural salience for each unique species by adding up the salience and dividing by the total number of participants in the study (n=59)
cultural_salience_df <- SC %>%
  group_by(species.preferred) %>%
  summarise(CS = sum(salience) / num_interviews_before_NA )

# Display the result
print(cultural_salience_df)
```

```{r}
# Reorder the data frame based on salience scores
cultural_salience_df <- cultural_salience_df[order(cultural_salience_df$CS, decreasing = TRUE), ]
# Create a new column for color based on the presence of 'shark' in the species name
cultural_salience_df$color <- ifelse(grepl("shark", tolower(cultural_salience_df$species.preferred)), "Sharks", "Rays")
```

```{r}
# Create a custom color palette
my_palette <- c("Sharks" = "#1f77b4", "Rays" = "#ff7f0e")

# Create the bar chart using ggplot2 with color based on species
CS_plot <- ggplot(cultural_salience_df, aes(x = reorder(species.preferred, -CS), y = CS, fill = color)) +
  geom_bar(stat = "identity") +
  labs(x = "Species",
       y = "Cultural Salience") +
  scale_fill_manual(values = my_palette, name = "Species", labels = c("Ray", "Shark")) +  # Use custom color palette with legend
  theme_minimal() +
  theme(axis.text.x = element_text(size = 8, angle = 45, hjust = 1)) +
  guides(fill = guide_legend(title = "Species"))

CS_plot
```

```{r}
# Calculating statistical test to test if the cultural salience of species are significant from each other
# Shapiro-Wilk test for normality
shapiro.test(cultural_salience_df$CS)
# The p-value is less than 0.05, so the data is not normally distributed and we will use the non-parametric Kruskal-Wallis test
```

```{r}
# Kruskal-Wallis test 
kruskal.test(CS ~ species.preferred, data = cultural_salience_df)
```

### Sankey Diagram for views on Protection

```{r}
library(htmltools)
library(highcharter)
library(htmlwidgets)
```

```{r}
sankey.data1.0 <- read.csv("views.csv")
```

```{r}
head(sankey.data1.0)
```

```{r}
# First Sankey diagram
s1 <- hchart(data_to_sankey(sankey.data1.0), "sankey", name = "Fish Type and Views on Protection")
```

```{r}
# Add title and caption
s1 <- s1 %>% hc_title(text = "Fish Type and Location based Views on Protection", align = "center")
s1
```

## Adding Value labels and percentages to the Sankey 
```{r}
# Get the count of the fish type and create new column which 
# has the name as well as the count in it
sankey.data2.0 <- sankey.data1.0 %>%
  dplyr::group_by(Fish.Type)%>%
      dplyr::tally()%>%
      dplyr::mutate(perc = n/sum(n))%>%
dplyr::mutate(Fish.TypelNew = paste(Fish.Type, n, '(', round(perc* 100,1) , '%)'))%>%
dplyr::select(-n, - perc)
sankey.data2.0
```

```{r}
sankeyMain <- merge (sankey.data1.0, sankey.data2.0, by = "Fish.Type")

# Get the count of the Market and create new column which 
# has the Market as well as the count in it

sankey.data3.0 <- sankeyMain %>%
  dplyr::group_by(Market)%>%
  dplyr::tally()%>%
  dplyr::mutate(perc = n/sum(n))%>%
  dplyr::mutate(MarketNew = paste(Market, n, '(', round(perc* 100,1) , '%)'))%>%
  dplyr::select(-n, - perc)

sankeyMain <- merge (sankeyMain, sankey.data3.0, by = "Market")
```

```{r}
# Get the count of the Views and create new column which
# has the Views as well as the count in it

sankey.data4.0 <- sankeyMain %>%
  dplyr::group_by(Views)%>%
  dplyr::tally()%>%
  dplyr::mutate(perc = n/sum(n))%>%
  dplyr::mutate(ViewsNew = paste(Views, n, '(', round(perc* 100,1) , '%)'))%>%
  dplyr::select(-n, - perc)

sankeyMain <- merge (sankeyMain, sankey.data4.0, by = "Views")
```

```{r}
sankeyFinal <- sankeyMain %>%
  dplyr::select(Fish.TypelNew, MarketNew, ViewsNew)

sankeyFinal
```

```{r}
library(highcharter)

# Define colors
publication_colors <- c("#1f77b4", "#d62728", "#2ca02c", "#ff7f0e", "#bcbd22")

# Create the Sankey diagram
s2 <- hchart(data_to_sankey(sankeyFinal), "sankey", name = "Fish Type and Location based Views on Protection") %>%
  hc_title(text= "Fish Type and Location based Views on Protection") %>%
  hc_colors(c(publication_colors))  # Set colors

s2

```


```{r}
# save the sankey diagram as an image
saveWidget(s2, "sankey.html", selfcontained = TRUE)
```

# Sankey diagaram for the restaurant data views
```{r}
sankey.data2.0 <- read.csv("views.restaurant.csv")
```

```{r}
# Draw the sankey diagram
# First Sankey diagram
s2.1 <- hchart(data_to_sankey(sankey.data2.0), "sankey", name = "Restaurant Views on Protection")
s2.1
```
## Chi squared test 

```{r}
sankey.chi <- read.csv("views.age.csv")
print(sankey.chi)
```

```{r}
# Define the age categories
age_categories <- c('Below 30', '30-40', '40-50', '50-60', '60+')

# Create a new column 'Age.Category' based on 'Age'
sankey.chi$Age.Category <- cut(sankey.chi$Age, breaks = c(-Inf, 30, 40, 50, 60, Inf), labels = age_categories)
```


```{r}
# Convert 'Age.Category' to a factor with the specified levels
sankey.chi$Age.Category <- factor(sankey.chi$Age.Category, levels = age_categories)

# Check the structure of the data frame to confirm the changes
str(sankey.chi)
```

```{r}
# Create the contingency table and perform the chi-squared test
contingency_table1 <- table(sankey.chi$Views, sankey.chi$Age.Category)
chi_sq_result <- chisq.test(contingency_table1)

# Print the result
print(chi_sq_result)

#Not significant (p=0.469)
```

```{r}
# Create a contingency table for 'Market' and 'Views'
contingency_table_market_views <- table(sankey.chi$Market, sankey.chi$Views)

# Perform chi-squared test
chi_sq_result_market_views <- chisq.test(contingency_table_market_views)

# Print the result
print(chi_sq_result_market_views)

#Not significant (p=0.3455)
```

```{r}
contingency_table_market <- table(sankey.chi$Fish.Type, sankey.chi$Views)

# Perform chi-squared test
chi_sq_result_market<- chisq.test(contingency_table_market)

# Print the result
print(chi_sq_result_market)

#Not significant (p=0.0.4473)
```




